---
title: "02DataModelling"
author: "FR"
date: "05/03/2021"
output: html_document
---

## Setup data paths
```{r}
expPath <- "expResults/exAuto1604/"
curveDf <- data.frame("batchSize" = c(8,4,2,1), "s" = c(1,1.7,3,5.3))
nCores <- 10

library(SPOT)
library(flacco)
library(dplyr)
library(randomForest)
require(pbapply)
require(pbmcapply)
```


## Gathering the points from evaluated algo runs
```{r}
if(!file.exists("expResults/exdataWithX.rds")){
    source("helperScriptsFunctions/readBBOB.R")
    source("helperScriptsFunctions/bbobResultAnalysis.R")
    
    df <- readBBOB(expPath)
    df <- applyFrNamingScheme(df, keepX = TRUE, keepY = TRUE)
    df <- df %>% select(-"seed")
    df$realY <- df$`df[, startsWith(names(df), "measuredFitnes")]`
    df <- df %>% select(-'df[, startsWith(names(df), "measuredFitnes")]')
    
    variablesInAlgoName = c("seed","functionID","algoID","nDim","budget","batchSize")
    algorithmIDsToReplace = c(1,2,3,4,5,6,7,8,9,10,12,13)
    algorithmNamesToInsert = c("RandomSearch",
                               "SPOT",
                               "DIRECT",
                               "1+1ES",
                               "moimbo",
                               "Q-EI",
                               "IPI",
                               "CMAES",
                               "DE",
                               "BOBYQA",
                               "NEWUOA",
                               "MultiLocalEI")
    
    df <- appendVariablesByAlgoName(df, variablesInAlgoName)
    df$algName <- replaceConfigVector(df$algoID, algorithmIDsToReplace, algorithmNamesToInsert)
    df$nDim <- as.numeric(as.character(df$nDim))
    df <- unique(df)
    
    saveRDS(df,file = "expResults/exdataWithX.rds")
}
df <- readRDS("expResults/exdataWithX.rds")
```

## Preparing data removing nas etc
```{r}
removeNACols <- function(df){
    df[, colSums(is.na(df)) != nrow(df)]
}

df <- df %>% filter(algName == "MultiLocalEI")
df <- df %>% filter(batchSize != "1")
df <- removeNACols(df)

df$iteration <- trunc((df$iteration-1) / as.numeric(df$batchSize)) + 1
df$unitTime <- as.numeric(df$iteration)
df$batchSize <- as.numeric(df$batchSize)
for(bs in curveDf$batchSize){
    df$unitTime[df$batchSize == bs] <- df$unitTime[df$batchSize == bs] / curveDf$s[curveDf$batchSize == bs]
}

df <- df[df$unitTime <= 50, ]
df <- df[df$batchSize != 4,]

getBestConfigBySeed <- function(g,...){
    filtered <- g %>% group_by(batchSize) %>% slice(tail(row_number(), 1)) %>% 
        group_by(batchSize) %>% mutate(med = median(y))
    g$bestBatchSizeBySeed <- as.numeric(max(filtered$batchSize[filtered$med == min(filtered$med)]))
    return(g)
}

getBestConfig <- function(g,...){
    filtered <- g %>% group_by(batchSize,seed) %>% slice(tail(row_number(), 1)) %>% 
        group_by(batchSize) %>% mutate(med = median(y))
    #g$bestBatchSize <- as.numeric(max(filtered$batchSize[filtered$med == min(filtered$med)]))
    g$bestBatchSize <- mean(g$bestBatchSizeBySeed)
    return(g)
}

getBestMedianConfig <- function(g,...){
    filtered <- g %>% group_by(batchSize,seed) %>% slice(tail(row_number(), 1)) %>% 
        group_by(batchSize) %>% mutate(med = median(y))
    #g$bestBatchSize <- as.numeric(max(filtered$batchSize[filtered$med == min(filtered$med)]))
    g$bestMedianBatchSize <- as.numeric(max(filtered$batchSize[filtered$med == min(filtered$med)]))
    return(g)
}

df <- df %>% group_by(functionID, nDim, algName, algoID) %>% group_modify(getBestMedianConfig)
df <- df %>% group_by(functionID, nDim, algName, algoID, seed) %>% group_modify(getBestConfigBySeed)
df <- df %>% group_by(functionID, nDim, algName, algoID) %>% group_modify(getBestConfig)

df$functionID <- as.numeric(as.character(df$functionID))
```

## calculating the features on that data
```{r, echo=FALSE, warning=FALSE}
getFeatures <- function(g, ...){
    gX <- removeNACols(g[,startsWith(names(g),"x")])
    gX <- as.matrix(gX)
    fObj <- createFeatureObject(X = gX, y = g$realY)
    ctrl <- list(allow_cellmapping = FALSE, show_progress = F, 
                             subset = c("basic","ela_meta",
                                        "ic"))
    features <- data.frame(calculateFeatures(fObj, control = ctrl))
    features$functionID = g$functionID[1]
    features$seed = g$seed[1]
    features$nDim <- g$nDim[1]
    return(features)
}

if(!file.exists("expResults/featDF.rds")){
    featDF <- df %>% group_by(functionID, nDim, algName, algoID, batchSize, seed) %>% group_modify(getFeatures)
    saveRDS(featDF, file = "expResults/featDF.rds")
}

featDF <- readRDS("expResults/featDF.rds")
```

## Build a data set with features and prediction results
```{r}
prepareAnalysisDf <- function(df, featDF){
    sliceDf <- df %>% group_by(functionID, nDim, algName, algoID, batchSize, seed) %>% slice(tail(row_number(), 1))
    sliceDf <- sliceDf[,-which(startsWith(names(sliceDf),"x"))]
    sliceDf <- sliceDf[, -which(names(sliceDf) %in% c("y", "iteration", "unitTime", "budget"))]
    featDF$functionID <- as.numeric(as.character(featDF$functionID))
    
    analysisDf <- full_join(sliceDf, featDF)
    analysisDf <- analysisDf[, -c(12:25)]
    
    completeAnalysisDF <- do.call(data.frame,lapply(analysisDf, function(x) replace(x, is.infinite(x),NA)))
    completeAnalysisDF <- completeAnalysisDF[complete.cases(completeAnalysisDF),]
    completeAnalysisDF$bestMedianBatchSize <- as.factor(completeAnalysisDF$bestMedianBatchSize)
    completeAnalysisDF$seed <- as.numeric(as.character(completeAnalysisDF$seed))
    completeAnalysisDF <- completeAnalysisDF[,!grepl("costs",names(completeAnalysisDF))]
    return(completeAnalysisDF)
}

completeAnalysisDF <- prepareAnalysisDf(df, featDF)
saveRDS(completeAnalysisDF,"expResults/completeAnalysisDF.rds")
```

## Feature selection with caret
```{r}
# ensure the results are repeatable
set.seed(7)
# load the library
library(mlbench)
library(caret)

# load the data
X <- completeAnalysisDF[,11:25]
y <- completeAnalysisDF$bestMedianBatchSize

index <- list()
xFunID <- completeAnalysisDF$functionID
for(i in 1:24){
    trainFuns <- c(1:24)[-i]
    trainInd <- which(xFunID %in% trainFuns)
    index <- c(index, list(trainInd))
}

unique(completeAnalysisDF[index[[1]],]$seed)

library(doParallel)
cl <- makePSOCKcluster(10)
registerDoParallel(cl)

# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", index = index)

# run the RFE algorithm
results <- rfe(X, y, sizes=c(1:10), rfeControl=control, metric = "Kappa")

stopCluster(cl)

# summarize the results
print(results)
# list the chosen features
predictors(results)
# plot the results
plot(results, type=c("g", "o"))
```
## Get Random Accuracy for Kappa value
```{r}
randAcc <- NULL

for(i in 1:10){
    Xsub <- completeAnalysisDF[, which(names(completeAnalysisDF) %in% predictors(results))]
    iid <- completeAnalysisDF$seed
    y <- completeAnalysisDF$bestMedianBatchSize
    
    trainInd <- sample(unique(iid), size = 0.75*length(unique(iid)), replace = F)
    testInd <- (1:max(unique(iid)))[-trainInd]
    
    rf <- randomForest(x = Xsub[completeAnalysisDF$seed %in% trainInd,], 
                       y = y[completeAnalysisDF$seed %in% trainInd], 
                       na.action = na.omit)
    
    y <- sample(y)
    
    yPred <- predict(rf, Xsub[completeAnalysisDF$seed %in% testInd,])
    lAcc <- sum(yPred == y[completeAnalysisDF$seed %in% testInd])/sum(completeAnalysisDF$seed %in% testInd)
    randAcc <- c(randAcc, lAcc)
}
print(randAcc)
randAcc <- mean(randAcc)
print(randAcc)

getKappa <- function(p, pc){
    return(
        (p-pc) / (1-pc)
    )
}
```



## Build and test a random forest with the proposed predictors (Level 1 - easy - completly random selection)
```{r}
Xsub <- completeAnalysisDF[, which(names(completeAnalysisDF) %in% predictors(results))]
iid <- completeAnalysisDF$seed
y <- completeAnalysisDF$bestMedianBatchSize

trainInd <- sample(unique(iid), size = 0.75*length(unique(iid)), replace = F)
testInd <- (1:max(unique(iid)))[-trainInd]
#trainInd <- completeAnalysisDF$seed %in% trainInd
#testInd <- completeAnalysisDF$seed %in% testInd

rf <- randomForest(x = Xsub[completeAnalysisDF$seed %in% trainInd,], 
                   y = y[completeAnalysisDF$seed %in% trainInd], 
                   na.action = na.omit)
rf

yPred <- predict(rf, Xsub[completeAnalysisDF$seed %in% testInd,])
testAcc <- sum(yPred == y[completeAnalysisDF$seed %in% testInd])/sum(completeAnalysisDF$seed %in% testInd)
print(paste("Test accuracy:", testAcc))
print(paste("Test kappa:", getKappa(testAcc, randAcc)))
```

## Build and save a model for later use
```{r}
trainCols <- names(Xsub)
rf <- randomForest(x = Xsub, 
                   y = y, 
                   na.action = na.omit)

saveRDS(object = list("rf"=rf, "trainCols"=trainCols), file = "expResults/autoRF.rds")
```

## Simulation Setup
```{r}
expPath <- "expResults/exAuto2704/"
```

## Gathering the points from evaluated algo runs

```{r setup, include=FALSE}
    source("helperScriptsFunctions/readBBOB.R")
    source("helperScriptsFunctions/bbobResultAnalysis.R")
    
    df <- readBBOB(expPath, filterString = "13Multi")
    df$funID <- as.numeric(as.character(df$funID))
    df <- df[df$funID > 24,]
    
    df <- applyFrNamingScheme(df, keepX = TRUE, keepY = TRUE)
    df <- df %>% select(-"seed")
    df$realY <- df$`df[, startsWith(names(df), "measuredFitnes")]`
    df <- df %>% select(-'df[, startsWith(names(df), "measuredFitnes")]')
    
    variablesInAlgoName = c("seed","functionID","algoID","nDim","budget","batchSize")
    algorithmIDsToReplace = c(1,2,3,4,5,6,7,8,9,10,12,13,14)
    algorithmNamesToInsert = c("RandomSearch",
                               "SPOT",
                               "DIRECT",
                               "1+1ES",
                               "moimbo",
                               "Q-EI",
                               "IPI",
                               "CMAES",
                               "DE",
                               "BOBYQA",
                               "NEWUOA",
                               "MultiLocalEI",
                               "AutoMultiLocalEI")
    
    df <- appendVariablesByAlgoName(df, variablesInAlgoName)
    df$algName <- replaceConfigVector(df$algoID, algorithmIDsToReplace, algorithmNamesToInsert)
    df$nDim <- as.numeric(as.character(df$nDim))
    df <- unique(df)
```


## Preparing data removing nas etc
```{r}
removeNACols <- function(df){
    df[, colSums(is.na(df)) != nrow(df)]
}

df <- df %>% filter(algName == "MultiLocalEI")
df <- df %>% filter(batchSize != "1")
df <- removeNACols(df)

df$iteration <- trunc((df$iteration-1) / as.numeric(df$batchSize)) + 1
df$unitTime <- as.numeric(df$iteration)
df$batchSize <- as.numeric(df$batchSize)
for(bs in curveDf$batchSize){
    df$unitTime[df$batchSize == bs] <- df$unitTime[df$batchSize == bs] / curveDf$s[curveDf$batchSize == bs]
}

df <- df[df$unitTime <= 50, ]
df <- df[df$batchSize != 4,]

getBestMedianConfig <- function(g,...){
    filtered <- g %>% group_by(batchSize,seed) %>% slice(tail(row_number(), 1)) %>% 
        group_by(batchSize) %>% mutate(med = median(y))
    #g$bestBatchSize <- as.numeric(max(filtered$batchSize[filtered$med == min(filtered$med)]))
    g$bestMedianBatchSize <- as.numeric(max(filtered$batchSize[filtered$med == min(filtered$med)]))
    return(g)
}

df <- df %>% group_by(functionID, nDim, algName, algoID) %>% group_modify(getBestMedianConfig)

getBestConfigBySeed <- function(g,...){
    filtered <- g %>% group_by(batchSize) %>% slice(tail(row_number(), 1)) %>% 
        group_by(batchSize) %>% mutate(med = median(y))
    g$bestBatchSizeBySeed <- as.numeric(max(filtered$batchSize[filtered$med == min(filtered$med)]))
    return(g)
}

df <- df %>% group_by(functionID, nDim, algName, algoID, seed) %>% group_modify(getBestConfigBySeed)


df$functionID <- as.numeric(as.character(df$functionID))
```

## calculating the features on that data
```{r, echo=FALSE, warning=FALSE}
getFeatures <- function(g, ...){
    gX <- removeNACols(g[,startsWith(names(g),"x")])
    gX <- as.matrix(gX)
    fObj <- createFeatureObject(X = gX, y = g$realY)
    ctrl <- list(allow_cellmapping = FALSE, show_progress = F, 
                             subset = c("basic","ela_meta",
                                        "ic"))
    features <- data.frame(calculateFeatures(fObj, control = ctrl))
    features$functionID = g$functionID[1]
    features$seed = g$seed[1]
    features$nDim <- g$nDim[1]
    return(features)
}

groupList <- df %>% group_by(functionID, nDim, algName, algoID, batchSize, seed) %>% group_split()
featDF <- bind_rows(pbmcapply::pbmclapply(groupList, getFeatures, mc.cores = nCores)) 

```

## Build a data set with features and prediction results
```{r}
prepareAnalysisDf <- function(df, featDF){
    sliceDf <- df %>% group_by(functionID, nDim, algName, algoID, batchSize, seed) %>% slice(tail(row_number(), 1))
    sliceDf <- sliceDf[,-which(startsWith(names(sliceDf),"x"))]
    sliceDf <- sliceDf[, -which(names(sliceDf) %in% c("y", "iteration", "unitTime", "budget"))]
    featDF$functionID <- as.numeric(as.character(featDF$functionID))
    
    analysisDf <- full_join(sliceDf, featDF)
    analysisDf <- analysisDf[, -c(10:23)]
    
    completeAnalysisDF <- do.call(data.frame,lapply(analysisDf, function(x) replace(x, is.infinite(x),NA)))
    completeAnalysisDF <- completeAnalysisDF[complete.cases(completeAnalysisDF),]
    completeAnalysisDF$bestMedianBatchSize <- as.factor(completeAnalysisDF$bestMedianBatchSize)
    completeAnalysisDF$seed <- as.numeric(as.character(completeAnalysisDF$seed))
    completeAnalysisDF <- completeAnalysisDF[,!grepl("costs",names(completeAnalysisDF))]
    return(completeAnalysisDF)
}
df$functionID <- as.numeric(as.character(df$functionID))
completeAnalysisDF <- prepareAnalysisDf(df, featDF)
saveRDS(completeAnalysisDF,"expResults/completeSimulatedAnalysisDF.rds")
```