## used

---
title: "03FeatureTests"
author: "FR"
date: "29/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(smoof)
require(dplyr)
require(multidplyr)
require(parallel)
require(pbapply)
require(pbmcapply)
library(flacco)

# load the library
library(mlbench)
library(caret)
library(doParallel)
```

## This document tries to validate the feature based prediction based on some randomly generated points evaluated on BBOB
For this purpose, 500 uniform random samples are evaluated on the first 15 instances of each BBOB function.
This data is then used in some model fitting experiments. 


### Generating a lot of samples to generate the features off
```{r}
set.seed(7)
nDim = 5
nSamples <- 500
nCores <- 10

X <- matrix(runif(nSamples*nDim,min = -5, max = 5), ncol = 5)
listOfData <- list()
for(i in 1:24){
    for(j in 1:15){
        fun <- smoof::makeBBOBFunction(dimensions = nDim, fid = i, iid = j)
        results <- fun(t(X))
        listOfData[[length(listOfData) + 1]] <- data.frame(X, results, nDim, "fid" = i, iid = j, rowID = 1:nSamples)
    }
}
resultDf <- bind_rows(listOfData)
rm(listOfData)
```

### Calculating the feature set on the collected data
```{r}
removeNACols <- function(df){
    df[, colSums(is.na(df)) != nrow(df)]
}

getFeatures <- function(g, ...){
    gX <- removeNACols(g[,startsWith(names(g),"X")])
    gX <- as.matrix(gX)
    fObj <- createFeatureObject(X = gX, y = g$results)
    ctrl <- list(allow_cellmapping = FALSE, show_progress = F, 
                             subset = c("basic","ela_meta",
                                        "ic"))
    features <- data.frame(calculateFeatures(fObj, control = ctrl))
    features$nDim <- g$nDim[1]
        features$fid <- g$fid[1]
        features$iid <- g$iid[1]
    return(features)
}

groupList <- resultDf %>% group_by(fid, nDim, iid) %>% group_split()
featureDF <- bind_rows(pbmcapply::pbmclapply(groupList, getFeatures, mc.cores = nCores))
```

## Add the function group and create completeAnalysisDF for prediction
```{r}
completeAnalysisDF <- featureDF
completeAnalysisDF$funGroup <- ""
completeAnalysisDF$funGroup[completeAnalysisDF$fid < 6] <- "Separable"
completeAnalysisDF$funGroup[completeAnalysisDF$fid < 10 & completeAnalysisDF$fid > 5] <- "lowCond"
completeAnalysisDF$funGroup[completeAnalysisDF$fid < 15 & completeAnalysisDF$fid > 9] <- "highCond"
completeAnalysisDF$funGroup[completeAnalysisDF$fid < 20 & completeAnalysisDF$fid > 14] <- "adeqGlobalStruct"
completeAnalysisDF$funGroup[completeAnalysisDF$fid > 19] <- "weakGlobalStruct"

completeAnalysisDF <- do.call(data.frame,lapply(completeAnalysisDF, function(x) replace(x, is.infinite(x),NA)))
completeAnalysisDF <- completeAnalysisDF[complete.cases(completeAnalysisDF),]
completeAnalysisDF <- completeAnalysisDF[,!grepl("costs",names(completeAnalysisDF))]
```

## Find the best predictors by rfe
```{r}
# ensure the results are repeatable
set.seed(5)


# load the data
X <- completeAnalysisDF[,c(14:28)]
y <- as.factor(completeAnalysisDF$funGroup)

index <- list()
xFunID <- as.numeric(as.character(completeAnalysisDF$iid))
for(i in 1:15){
    trainFuns <- c(1:15)[-i]
    trainInd <- which(xFunID %in% trainFuns)
    index <- c(index, list(trainInd))
}


cl <- makePSOCKcluster(10)
registerDoParallel(cl)

# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv")#, index = index, verbose = T)

# run the RFE algorithm
results <- rfe(X, y, sizes=c(1:10), rfeControl=control, metric = "Kappa")

stopCluster(cl)

# summarize the results
print(results)
# list the chosen features
predictors(results)
# plot the results
plot(results, type=c("g", "o"))
```
## Try to measure some distances between instances and between functions
```{r}
chosenPredictors <- c(predictors(results)[1:5],"fid")
#X <- X[, which(names(X) %in% chosenPredictors)]
#X <- BBmisc::normalize(X)#, method = "range") ## best normalization method should be considered?

## function to calculate a summary on distances based on a given feature DF
summaryDistDF <- function(featDf){
    featDf <- featDf[,-which(names(featDf) == "fid")]
    s <- summary(dist(featDf))
    df <- t(data.frame(x=matrix(s),row.names=names(s)))
    rownames(df) <- NULL
    return(df)
}

distanceData <- completeAnalysisDF[,which(names(completeAnalysisDF) %in% chosenPredictors)]
distanceData[,1:5] <- BBmisc::normalize(distanceData[,1:5])

distDF <- NULL
for(fid in unique(completeAnalysisDF$fid)){
    subData <- distanceData[distanceData$fid==fid,]
    distDF <- rbind(distDF, cbind(data.frame("fid" = fid), summaryDistDF(subData)))
}

distDF <- rbind(distDF, cbind(data.frame("fid" = 0), summaryDistDF(distanceData)))
tail(distDF)
```

## Try to measure some distances between instances and between functions
```{r}
chosenPredictors <- c(predictors(results)[1:5],"funGroup")
#X <- X[, which(names(X) %in% chosenPredictors)]
#X <- BBmisc::normalize(X)#, method = "range") ## best normalization method should be considered?

## function to calculate a summary on distances based on a given feature DF
summaryDistDF <- function(featDf){
    featDf <- featDf[,-which(names(featDf) == "funGroup")]
    s <- summary(dist(featDf))
    df <- t(data.frame(x=matrix(s),row.names=names(s)))
    rownames(df) <- NULL
    return(df)
}

distanceData <- completeAnalysisDF[,which(names(completeAnalysisDF) %in% chosenPredictors)]
distanceData[,1:5] <- BBmisc::normalize(distanceData[,1:5])

distDF <- NULL
for(funGroup in unique(completeAnalysisDF$funGroup)){
    subData <- distanceData[distanceData$funGroup==funGroup,]
    distDF <- rbind(distDF, cbind(data.frame("funGroup" = funGroup), summaryDistDF(subData)))
}

distDF <- rbind(distDF, cbind(data.frame("funGroup" = 0), summaryDistDF(distanceData)))
tail(distDF)
```

